---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HDNRA

<!-- badges: start -->

[![R-CMD-check](https://github.com/nie23wp8738/HDNRA/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nie23wp8738/HDNRA/actions/workflows/R-CMD-check.yaml) [![License: GPL-3.0](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)

<!-- badges: end -->

The R package **HDNRA** includes the latest methods based on normal-reference approach to test the equality of the mean vectors of high-dimensional samples with possibly different covariance matrices. `HDNRA` is also used to demonstrate the implementation of these tests, catering not only to the two-sample problem, but also to the general linear hypothesis testing (GLHT) problem. This package provides easy and user-friendly access to these tests. Both coded in C++ to allow for reasonable execution time using [Rcpp](https://github.com/RcppCore/Rcpp). Besides [Rcpp](https://github.com/RcppCore/Rcpp), the package has no strict dependencies in order to provide a stable self-contained toolbox that invites re-use.

There are two real data sets in `HDNRA`: [COVID19](https://nie23wp8738.github.io/HDNRA/reference/COVID19.html) and [corneal](https://nie23wp8738.github.io/HDNRA/reference/corneal.html).

Seven normal-reference tests for the two-sample problem: [ts_zgzc2020()](https://nie23wp8738.github.io/HDNRA/reference/ts_zgzc2020.html), [ts_zz2022()](https://nie23wp8738.github.io/HDNRA/reference/ts_zz2022.html), [ts_zzz2020()](https://nie23wp8738.github.io/HDNRA/reference/ts_zzz2020.html), [tsbf_zwz2023()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_zwz2023.html), [tsbf_zz2022()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_zz2022.html), [tsbf_zzgz2021()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_zzgz2021.html), [tsbf_zzz2023()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_zzz2023.html).

Five normal-reference tests for the GLHT problem in `HDNRA`: [glht_zgz2017()](https://nie23wp8738.github.io/HDNRA/reference/glht_zgz2017.html), [glht_zz2022()](https://nie23wp8738.github.io/HDNRA/reference/glht_zz2022.html), [glht_zzz2022()](https://nie23wp8738.github.io/HDNRA/reference/glht_zzz2022.html), [glhtbf_zz2022()](https://nie23wp8738.github.io/HDNRA/reference/glhtbf_zz2022.html), [glhtbf_zzg2022()](https://nie23wp8738.github.io/HDNRA/reference/glhtbf_zzg2022.html).

Four existing tests for the two-sample problem in `HDNRA`: [ts_bs1996()](https://nie23wp8738.github.io/HDNRA/reference/ts_bs1996.html), [ts_sd2008()](https://nie23wp8738.github.io/HDNRA/reference/ts_sd2008.html), [tsbf_cq2010()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_cq2010.html), [tsbf_skk2013()](https://nie23wp8738.github.io/HDNRA/reference/tsbf_skk2013.html).

Five existing tests for the GLHT problem in `HDNRA`: [glht_fhw2004()](https://nie23wp8738.github.io/HDNRA/reference/glht_fhw2004.html), [glht_sf2006()](https://nie23wp8738.github.io/HDNRA/reference/glht_sf2006.html), [glht_ys2012()](https://nie23wp8738.github.io/HDNRA/reference/glht_ys2012.html), [glhtbf_zgz2017()](https://nie23wp8738.github.io/HDNRA/reference/glhtbf_zgz2017.html), [ks_s2007()](https://nie23wp8738.github.io/HDNRA/reference/ks_s2007.html).

## Installation

You can install and load the most recent development version of `HDNRA` from [GitHub](https://github.com/) with:

``` r
# Installing from GitHub requires you first install the devtools or remotes package
install.packages("devtools")
# Or
install.packages("remotes")

# install the most recent development version from GitHub
devtools::install_github("nie23wp8738/HDNRA")
# Or
remotes::install_github("nie23wp8738/HDNRA")
# load the most recent development version from GitHub
library(HDNRA)
```

## Usage

### Load the package

```{r}
library(HDNRA)
```

### Example data

Package `HDNRA` comes with two real data sets:

```{r}
# A COVID19 data set from NCBI with ID GSE152641.
?COVID19

# A corneal data set acquired during a keratoconus study.
?corneal
```

### Example for two-sample problem

A simple example of how to use one of the normal-reference tests `tsbf_zwz2023` using data set `COVID19`:

```{r}
data("COVID19")
group1 <- as.matrix(COVID19[c(2:19, 82:87), ]) # healthy group1
group2 <- as.matrix(COVID19[-c(1:19, 82:87), ]) # patients group2
# The data matrix for tsbf_zwz2023 should be p by n, sometimes we should transpose the data matrix
tsbf_zwz2023(t(group1), t(group2))
```

### Example for GLHT problem

A simple example of how to use one of the normal-reference tests `glhtbf_zzg2022` using data set `corneal`:

```{r}
data("corneal")
p <- dim(corneal)[2]
k <- 4
Y <- list()
group1 <- as.matrix(corneal[1:43, ]) # normal
group2 <- as.matrix(corneal[44:57, ]) # unilateral suspect
group3 <- as.matrix(corneal[58:78, ]) # suspect
group4 <- as.matrix(corneal[79:150, ]) # clinical leratoconus
Y[[1]] <- t(group1)
Y[[2]] <- t(group2)
Y[[3]] <- t(group3)
Y[[4]] <- t(group4)
dim(Y[[1]])
dim(Y[[2]])
dim(Y[[3]])
dim(Y[[4]])
n <- c(43, 14, 21, 72)
G <- cbind(diag(k - 1), rep(-1, k - 1))
glhtbf_zzg2022(Y, G, n, p)
```

## Code of Conduct

Please note that the HDNRA project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
